LabMate: supporting types for MATLAB

### 2024-03-28 writing/IMEKOWC2024/LabMate/paper.tex
[ writing ] draft Implementation section

% transducer model

% elaboration into core type theory (with richer equational theory than state-of-the-art)

% Stack machine, with disjunctive problem solving

>>>>

In contrast to most other typecheckers, LabMate works using a
\remph{transducer} model of interaction: it inputs MATLAB code with
formal comments, and outputs a new version of its input, responding to
the comments, as if it were a development collaborator.  It thus needs
to make sure that it retains as much of the input formatting as
possible.  Making sense of MATLAB code presented a significant reverse
engineering challenge, as the MATLAB syntax is largely specified
informally and by example.

LabMate elaborates MATLAB expressions and commands into terms of its own internal core type theory.
%
Because we want to ergonomically handle physical dimensions, matrices
and their types, this type theory has a richer equational theory than
other state-of-the-art type theories, which we achieve by a refined
normalisation algorithm for free Abelian groups, and free monoids.
%
The actual typechecking is implemented as a stack based virtual machine which uses unification and a limited form of backtracking to solve typing constraints.
%
After the virtual machine has solved as many problems as it can, we analyse its final state to reconstruct the output source code, perhaps with inserted warnings and error messages, or additional responses to queries.

LabMate is under constant development; the latest version can be found at \url{https://github.com/msp-strath/LabMate/tree/devel}.
